---
title: "Various stock recruit relationships for depensatory dynamics"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, load pkgs, message = FALSE}
library(tidyverse)
```

## Background  
Here we explore various paramaterizations of stock recruit relationships (SRRs) that show evidence of depensation. There are many parameterizaitons for both Ricker and Beverton-Holt type SRRs but we will focus on some basics. Most of these derivations arise from a need for how to treat the depensatory parameter, with some having a goal of making the depensatory parameter "biologically relevant".    

### Ricker based SRRs  
There are 3 types we will explore here.  
- original [Ricker](https://cdnsciencepub.com/doi/10.1139/f54-039) from 1954   
- depensatory Ricker ([Quinn & Deriso, 1999](https://books.google.ca/books/about/Quantitative_Fish_Dynamics.html?id=5FVBj8jnh6sC&redir_esc=y), p. 99)  
- Saila-Lorda (as described in [Perala & Kuparinen, 2017](http://dx.doi.org/10.1098/rspb.2017.1284))  

To understand these models we'll describe their structure, simulate data based on the Saila-Lorda then fit each SRR to the simulated data.  

#### Original Ricker  
$$ R = \alpha Se^{-\beta S} $$  
where $\alpha$ is the slope near the origin and $\beta$ is the density dependent parameter. There is no term for density dependence in this equation, but we will use it to compare. The Ricker SRR is also popular because it can be transformed to be linear:  
$$ log(R/S) = \alpha - \beta S $$  

#### Quinn & Deriso's 4 parameter depensatory Ricker SRR  
$$ R = \alpha Se^{\beta S - \delta S^{\gamma }} $$
The authors say that this model is depensatory when $S^{\gamma -1} < \beta / (\gamma \delta)$, which isn't very intuitive, which brings us to the next model.    

#### Salia-Lorda SRR  
$$ R_{t} = k(\frac{S_{t}}{S_{k}})^{C}e^{C (1-S_{t}/S_{k})} $$
where $R_{t}$ and $S_{t}$ are vectors of recruits and spawners at time *t*, *k* is the maximum number of recruits produced when $S_{t} = S_{k}$ and $S_{k}$ is is the spawning stock at which the number of recruits is *k*. *c* is the depensatory parameter; *c* = 1 produces the classic Ricker, *c* > 1 shows depensatory dynamics, and *c* < 1 shows hypercompensaiton.  

## Simulation  
We'll simulate some data to play with. FEEL FREE to toggle the parameters (i.e. S, Sk, k, c, sigma), they will cascade through the robust code  

```{r, write fun}
sal_lor <- function(S, Sk, k, c, sigma){
  exp_rec <- ((k*(S/Sk)^c)*exp(c*(1-S/Sk))+sigma)
  return(exp_rec)
}
  
S <- runif(100, 10, 1000)
Sk <- 600
k <- 200
c <- 3
sigma <- rnorm(length(S), 0, 10)

exp_rec <- sal_lor(S, Sk, k, c, sigma)

plot(S, exp_rec)

sim_stock <- data.frame(stock = S, recruits = exp_rec)
```  

## Fit simulated data  
Can fit a Ricker here the "old fashioned" way by making it linear.  

```{r, ricker fit}
rick_fit <- lm(log(recruits/stock)~stock, data = sim_stock)

summary(rick_fit)$coefficients
```

now we can use optim to fit a non linear Quinn & Deriso model  
```{r, QD fit}
QDLogLike <- function(parms){
  alpha <- parms[1]
  beta <- parms[2]
  delta <- parms[3]
  gamma <- parms[4]
  sd <- exp(parms[5])
  
  predicted <- alpha*sim_stock$stock*exp((beta*sim_stock$stock)-(delta*sim_stock$stock^gamma))
  
  NegLogLike <- -sum(log(dnorm(sim_stock$recruits, predicted, sd= sd)))
  return(NegLogLike)
}

#declare start parms - guess based on previous fit 
alpha <- exp(summary(rick_fit)$coefficients[1])
beta <- -summary(rick_fit)$coefficients[2]
delta <- 1
gamma <- 1
sd <- log(10)

parms <- list(alpha, beta, delta, gamma, sd)

QD_fit <- optim(parms, QDLogLike, method="Nelder-Mead", hessian=TRUE)

QD_fit$par
```

Now we'll refit the Salia-Lorda  
```{r, sal-lor fit}
SLLogLike <- function(parms){
  Sk <- parms[1]
  k <- parms[2]
  c <- parms[3]
  sd <- exp(parms[4])
  
  predicted <- ((k*(sim_stock$stock/Sk)^c)*exp(c*(1-sim_stock$stock/Sk)))  
  NegLogLike <- -sum(log(dnorm(sim_stock$recruits, predicted, sd= sd)))
  return(NegLogLike)
}
#Sk <- 1000 #tester - this is finnecky! and changes results - must be hitting local minima
parms <- list(Sk, k, c, sd) #can recycle from above

SL_fit <- optim(parms, SLLogLike, method="Nelder-Mead", hessian=TRUE)
cbind(c("SK", "k", "c", "sd"), SL_fit$par)
```

Now we'll generate model predictions from the parameter fits   
```{r, plot it}
r_alpha <- exp(summary(rick_fit)$coefficients[1])
r_beta <- summary(rick_fit)$coefficients[2]

rick_preds <- r_alpha*sim_stock$stock*exp(r_beta*sim_stock$stock)
rick_preds <- data.frame(fit = rick_preds, S = S) %>%
  arrange(S)

QD_alpha <- QD_fit$par[1]
QD_beta <- QD_fit$par[2]
QD_delta <- QD_fit$par[3]
QD_gamma <- QD_fit$par[4]

QD_preds <- QD_alpha*sim_stock$stock*exp((QD_beta*sim_stock$stock)-(QD_delta*sim_stock$stock^QD_gamma))
QD_preds <- data.frame(fit = QD_preds, S = S) %>%
  arrange(S) 
SL_Sk <- SL_fit$par[1]
SL_k <- SL_fit$par[2]
SL_c <- SL_fit$par[3]

SL_preds <- ((SL_k*(sim_stock$stock/Sk)^c)*exp(c*(1-sim_stock$stock/Sk)))  

SL_preds <- data.frame(fit = SL_preds, S = S) %>%
  arrange(S)
```

Then plot it
```{r, plot fits}
plot(S, exp_rec)
lines(fit~S, data = rick_preds, col = "blue")
lines(fit~S, data = SL_preds, col = "green")
lines(fit~S, data = QD_preds, col = "red") #Something's messed up with the QD model!
```

## To-do  
- Clip the negative recruits out of the simulation!  
- Fix the 4 parameter Ricker from Quinn & Deriso so it actually fits!  
- Play around with different solvers (e.g. `nlme()`?) because `optim()` is very sensitive to starting parameters in these models  

we could add variations of the Beverton-Holt, which are generally easier to fit. These could include:  
- classic Beverton-Holt without depensation  
- Myers' Beverton-Holt from the '95 [paper](https://www.science.org/doi/10.1126/science.269.5227.1106) in *Science*  
- The Beverton-Holt fit in the Perala et al [paper](https://royalsocietypublishing.org/doi/10.1098/rsbl.2021.0439)  
- [Lierman & Hilborn's](https://cdnsciencepub.com/doi/abs/10.1139/f97-105) stepwise way of fitting a depensatory Beverton-Holt with meaningful parameters  
