---
title: "Simple depensatory models to all stocks"
output:
  html_document:
    collapsed: no
    fig_caption: yes
    smooth_scroll: yes
    theme: cerulean
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)

library(tidyverse)
library(ggh4x) #for some nicer facetting
library(here)
```

```{r, data}
SR_data <- read.csv(here("data/salmon_productivity_compilation_jun2022.csv")) %>%
  select(-X) %>%
  mutate(species_stock = paste(species, stock)) #to help with looping later

stock_info <- read.csv(here('data','all_stocks_info_jun2022.csv'))

```

# Fit Shepard models to all populaitons  
Need to get rid of some populations that won't fit.  
I think these aren't fitting because of high recruitment variability, which makes my simple init for the beta parm (i.e. `max(sub_stock$spawners)/1.5`) break the model.  
```{r, dump pops, echo = TRUE}
SR_data_filter <- filter(SR_data, !(species_stock %in% c("Chinook Andrew Creek", "Chinook Alsek/Klukshu",
                                                "Chinook Alsek-Klukshu",
                                                "Chinook East Fork South Fork Salmon River",
                                                "Chinook Entiat River", "Chinook Minam River",
                                                "Chinook Coweeman River"))) %>%
  mutate(spawners = ifelse(spawners==0|recruits==0, spawners+1, spawners),
         recruits = ifelse(spawners==0|recruits==0, recruits+1, recruits))
```

Then we can run the model  
```{r, shepard, echo = TRUE}
#build likliehood function
getNegLogLike <- function(parms){
  alpha <- parms[1]
  beta <- exp(parms[2])
  delta <- parms[3]
  sd <- exp(parms[4])
  
  predicted <- (alpha*sub_stock$spawners^delta)/(1+(beta*sub_stock$spawners^delta))
  NegLogLike <- -sum(log(dnorm(sub_stock$recruits, predicted, sd = sd)))
  return(NegLogLike)
}

#declare parms all pops will share
alpha <- 1
delta <- 1

shepard_results <- NULL #empty df to populate with results

for(i in unique(SR_data_filter$species_stock)){
  sub_stock <- filter(SR_data_filter, species_stock == i) %>%
      na.omit()
    
    #declare stock specific parms for beta and sd based on some heuristics 
    log_beta <- log(max(sub_stock$spawners)/1.5)
    log_sd <- log(sd(sub_stock$recruits/1.5)) 
    
    parms <- c(alpha, log_beta, delta, log_sd)
    
    #fit it
    fit1 <- optim(parms, getNegLogLike, method="Nelder-Mead", hessian=TRUE)
    fit2 <- optim(fit1$par, getNegLogLike, method="Nelder-Mead", hessian=TRUE)
    fit3 <- optim(fit2$par, getNegLogLike, method="Nelder-Mead", hessian=TRUE)
    

    shepard_results <- rbind(data.frame(i, unique(sub_stock$species), unique(sub_stock$stock),
                                        fit3$par[1], exp(fit3$par[2]), fit3$par[3], 
                                        exp(fit3$par[4]), unique(sub_stock$stock.id)),
                             shepard_results)
}
```

And then clean up the table  
```{r, shepard table clean}
colnames(shepard_results) <- c("species_stock", "species", "stock", "alpha", "beta", "delta", "sd", "stock.id")

shepard_results <- shepard_results %>% 
  mutate_at(4:7,funs(round(., 2)))
```

Some of these `beta` parameters are tiny. They must have had a hard time fitting. Let's look at the data to try and get an idea of whats going on. We'll also include the pops that wouldn't fit that we filtered out earlier.    
```{r, check bad stocks}
bad_stocks <- c(filter(shepard_results, beta<1) %>% pull(species_stock), 
                "Chinook Andrew Creek", "Chinook Alsek/Klukshu",
                                                "Chinook Alsek-Klukshu",
                                                "Chinook East Fork South Fork Salmon River",
                                                "Chinook Entiat River", "Chinook Minam River",
                                                "Chinook Coweeman River")


bad_stocks_data <- filter(SR_data, species_stock %in% bad_stocks)

max_facets <- 16

pages <- ceiling(length(bad_stocks)/max_facets)

for(i in 1:pages){
  sub_data <- filter(bad_stocks_data, species_stock %in% 
                               unique(bad_stocks_data$species_stock)[(((i-1)*max_facets)+1):(max_facets*i)])
  p <- ggplot(sub_data, aes(spawners, recruits)) +
    geom_point() +
        ggh4x::facet_wrap2(vars(species_stock), nrow=4, ncol=4, trim_blank=FALSE, 
                           scales="free")+
    labs(title = paste("Bad stocks", i, "of", pages))
  print(p)
}
```

## Plot delta

then let's take a look at the delta estimates by species, for fits we trust  
```{r, compare deltas, message = FALSE}
good_fits <- filter(shepard_results, !(species_stock %in% bad_stocks))

ggplot(good_fits, aes(delta)) +
  geom_histogram() +
  facet_wrap(~species) +
  geom_vline(xintercept=1, lty =2) +
  labs(title = "Histogram of delta parameter estimates form a depensatory beverton-holt")
```

```{r check sockeye, include = FALSE}
SK <- filter(shepard_results, species=="Sockeye")
data <- left_join(good_fits, stock_info, by = "stock.id") #UGH why u broken!?

```